Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-05 paper/Intuition.tex
Text cleanup

In this section, we present the structure of the proof of Theorem
\fullref{storeAtomicity} for a system implementing the MSI protocol of
Figure~\ref{trans}. We reduce the whole system proof to that of a set of
invariants between any parent-child pair. These invariants in turn depend on
exactly two invariants. Identifying these two basic invariants is one of the
main contributions of the paper; they also hold for other protocols like MOSI
(discussed in Section~\ref{Sec:Refinements}).

>>>>

In this section, we present the structure of the proof of Theorem
\fullref{storeAtomicity} for a system implementing the MSI protocol of
Figure~\ref{trans}. We reduce the whole system proof to that of a set
of invariants between any parent-child pair. These invariants in turn
depend on two smaller invariants which encapsulate the core aspects of
correctness of the MSI protocol and allow the large scope of verified
hardware refinements we present later (Section~\ref{Sec:Refinements}).

### 2014-02-05 paper/System.tex
Text cleanup

We have dedicated channels between each parent-child pair. The three channels
are indexed by the name of the child that uses that channel (denoted by
$\textit{ch}[c]$ for the \textit{ch} channel associated with cache $c$).
Figure \ref{format} shows the message format transmitted in the 3
channels between caches. In each of the channels $ch$, a message $m$
can be enqueued ($\enq(ch, m)$) or dequeued ($\deq(ch)$), and the
channels can be examined for the first element ($\first(ch)$) or for
availability of any message in the channel ($\avail(ch)$).  Each of
these channels start out empty \ie $\forall ch, \neg \avail(ch)$,
initially.

@@

\begin{subfigure}{\textwidth}
\centering
\begin{tabular}{|ll|}
\hline
\multicolumn{2}{|p{\textwidth}|}{\textbf{ChildVolResp}$(c, x, a)$: Child $c$
sending a response to downgrade to $x$ for address $a$ without any request from
its parent}\\
\hline
Guard: & $(\forall i, \parent(i, c) \rightarrow c.\dstate[i][a] \le x) \wedge x < c.\state[a] \wedge c.\wait[a] = \False$\\
\hline
Action: & $\enq(\cpResp[c], \langle c.\state[a], x, a,%$\\
%& $\;\;\;%
\myif c.\state[a] = \Mo \mythen c.\data[a] \myelse \_\rangle)$;\\
& $c.\state[a] \Leftarrow x$;\\
\hline
\hline
\multicolumn{2}{|p{\textwidth}|}{\textbf{ChildDropReq}$(c)$: Child $c$
receiving a request from its parent and has already downgraded}\\
\hline
Guard: & 
$\avail(\pc[c]) \wedge \mylet q := \first(\pc[c]) \myin q.\typ = \Req \wedge q.\myto \ge c.\state[q.\addr]$\\
\hline
Action: & $\deq(\pc[c])$;\\
\hline
\end{tabular}
\caption{Other transitions at a non-root cache}
\label{childextra}
\end{subfigure}

@@

Note that the atomic transitions in Figure~\ref{trans} access only local states,
\ie{} a transition can read or write states only corresponding to a single cache
and/or the channel the cache is connected to. This restriction allows such a
system to be implemented directly into hardware. Bluespec System Verilog (BSV),
for instance, directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV executes
several transitions simultaneously, as long as they do not access the same
state, \ie{} do not \emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
Sometimes, these transitions (both the guards and the actions) may not be
amenable to single cycle implementations in hardware. For example, writing a
register array can actually take several hardware clock cycles, though
logically it is a single transition.  Karczmarek \etal~\cite{Karczmarek} has
provided a scheme to convert atomic transitions into synchronous hardware in
which each transition can span several clock cycles.

>>>>

We have dedicated channels between each parent-child pair. The three
channels are indexed by the name of the child that uses that channel
(denoted by $\textit{ch}[c]$ for the \textit{ch} channel associated
with cache $c$).  Figure~\ref{format} shows the message format
transmitted in the 3 channels between caches. In each of the channels
$ch$, a message $m$ can be enqueued ($\enq(ch, m)$) or dequeued
($\deq(ch)$), and the channels can be examined for the first element
($\first(ch)$) or for availability of any message in the channel
($\avail(ch)$).  Each of these channels start out empty \ie{} $\forall
ch, \neg \avail(ch)$, initially.

@@

\begin{subfigure}{\textwidth}
\centering
\begin{tabular}{|ll|}
\hline
\multicolumn{2}{|p{\textwidth}|}{\textbf{ChildVolResp}$(c, x, a)$: Child $c$
sending a response to downgrade to $x$ for address $a$ without any request from
its parent}\\
\hline
Guard: & $(\forall i, \parent(i, c) \rightarrow c.\dstate[i][a] \le x) \wedge x < c.\state[a] \wedge c.\wait[a] = \False$\\
\hline
Action: & $\enq(\cpResp[c], \langle c.\state[a], x, a,
\myif c.\state[a] = \Mo \mythen c.\data[a] \myelse \_\rangle)$;\\
& $c.\state[a] \Leftarrow x$;\\
\hline
\hline
\multicolumn{2}{|p{\textwidth}|}{\textbf{ChildDropReq}$(c)$: Child $c$
receiving a request from its parent and has already downgraded}\\
\hline
Guard: & 
$\avail(\pc[c]) \wedge \mylet q := \first(\pc[c]) \myin q.\typ = \Req \wedge q.\myto \ge c.\state[q.\addr]$\\
\hline
Action: & $\deq(\pc[c])$;\\
\hline
\end{tabular}
\caption{Other transitions at a non-root cache}
\label{childextra}
\end{subfigure}

@@

Note that the atomic transitions in Figure~\ref{trans} access only
local states, \ie{} a transition can read or write states only
corresponding to a single cache and/or the channel the cache is
connected to. This restriction allows such a system to be implemented
directly into hardware. Bluespec SystemVerilog (BSV), for instance,
directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV
executes several transitions simultaneously in a single clock cycle,
as long as they do not access the same state, \ie{} do not
\emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.

%ndave: I know this is mildly relevant, but it's making the current
%hardware implementation seem worse and it's not adding any actual
%knowledge to the reader who doesn't care about this level of detail

%%  Sometimes,
%% these transitions may not be amenable to single cycle implementations
%% in hardware. For example, writing a memory can actually take several
%% hardware clock cycles, though logically it is a single transition.
%% Karczmarek \etal~\cite{Karczmarek} has provided a scheme to convert
%% atomic transitions into synchronous hardware in which each transition
%% can span several clock cycles.
